Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-25 paper.Rnw
updates based on comments from Heike - I'll work on proof reading next

# Load packages
library(caret)
library(cowplot)
library(dplyr)
library(forcats)
library(ggforce)
library(ggplot2)
library(ggpcp)
library(glmnet)
library(gower)
library(graphics)
library(gretchenalbrecht)
library(latex2exp)
library(lime) #devtools::install_github("goodekat/lime")
library(limeaid) #devtools::install_github("goodekat/limeaid")
library(purrr)
library(randomForest)
library(stringr)
library(tidyr)

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(data = lime_data_gathered,
             mapping = aes(x = feature_stnd_value,
                           y = prediction,
                           size = distance_value,
                           alpha = distance_value,
                           color = distance)) +
  geom_abline(data = coefs_data %>% filter(case == "good"),
              mapping = aes(intercept = int, slope = slope),
              size = 1) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c(NA, "grey30"), guide = FALSE) +
  geom_point(data = prediction_of_interest_gathered,
             mapping = aes(x = feature_value,
                           y = prediction,
                           fill = color),
             color = "black",
             size = 5,
             shape = 23,
             alpha = 0.75) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(feature_stnd_value = 1.2,
                      prediction = 1.6,
                      feature = c("Feature 1", "Feature 2"),
                      slope = c(paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "good",
                                               feature == "Feature 1") %>%
                                        pull(slope) %>%
                                        round(3)),
                                paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "good",
                                               feature == "Feature 2") %>%
                                        pull(slope) %>%
                                        round(3)))),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = 4) +
  labs(x = "",
       y = "Black-Box Prediction",
       title = "Conceptual Depiction of a Faithful Local Explainer Model",
       subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
       fill = "",
       alpha = "Weight",
       size = "Weight") +
  theme_linedraw(base_family = "Times", base_size = 10) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(color = 'black', size = 10)) +
  guides(fill = guide_legend(order = 1),
         size = guide_legend(order = 2),
         alpha = guide_legend(order = 2))

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(data = lime_data_gathered,
             mapping = aes(x = feature_stnd_value,
                           y = prediction,
                           size = distance_value,
                           alpha = distance_value,
                           color = distance)) +
  geom_abline(data = coefs_data %>% filter(case == "bad"),
              mapping = aes(intercept = int, slope = slope),
              size = 1) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c("grey30", NA), guide = FALSE) +
  geom_point(data = prediction_of_interest_gathered,
             mapping = aes(x = feature_value,
                           y = prediction,
                           fill = color),
             color = "black",
             size = 5,
             shape = 23,
             alpha = 0.75) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(feature_stnd_value = 1.2,
                      prediction = 1.6,
                      feature = c("Feature 1", "Feature 2"),
                      slope = c(paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "bad",
                                               feature == "Feature 1") %>%
                                        pull(slope) %>%
                                        round(3)),
                                paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "bad",
                                               feature == "Feature 2") %>%
                                        pull(slope) %>%
                                        round(3)))),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = 4) +
  labs(x = "",
       y = "Black-Box Prediction",
       title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
       subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
       fill = "",
       alpha = "Weight",
       size = "Weight") +
  theme_linedraw(base_family = "Times", base_size = 10) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(color = 'black', size = 10)) +
  guides(fill = guide_legend(order = 1),
         size = guide_legend(order = 2),
         alpha = guide_legend(order = 2))

@@

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two predictor variables used to fit the hypothetical model. The diamond shaped points represent a prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. \hh{In the example of \autoref{fig:concept-plot-good}} an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of \hh{two features in the data} % features Features 1 and~2 
as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

@@

To demonstrate the visualizations, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with three features and one response variable. The features, $x_1$, $x_2$, and $x_3$, are randomly sampled from Unif(\Sexpr{min}, \Sexpr{max}), Unif(\Sexpr{min}, \Sexpr{max}), and $\mbox{N}(0,1)$ distributions, respectively. A binary response variable $y$ is  created using a rotated sine curve. In particular, let $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$ and $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$ where $\theta=-0.9$. Then $y$ is defined as
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ . %\nolabel
  \end{cases}
\end{eqnarray}
Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$.
\hh{From the plots in \autoref{sine-plot-data} we see that the global relationship between response $y$ and features $x_1$ and $x_2$ is linear: the probability for label blue increases with the difference between features $x_2$ and $x_1$. Locally, the relationship between $y$ and features $x_2$ and $x_1$ varies a lot more around the line of identity. Here, the relationship  is determined by the sine waves. However, the sine is a good-natured function that can be approximated well linearly. }

@@

\hh{All runs of LIME in this paper are executed in}
a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. 
The forked version \hh{is functionally indistinguishable from Pedersen's implementation but it} allows us to export internal values \hh{relevant for an assessment of the explainer}. 
The R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid} (\Sexpr{packageVersion("limeaid")}) implements diagnostic plots based on the output of the altered version of \emph{lime}.

@@

\begin{figure*}[!thp]
<<sine-exp-scatter, out.width = '\\textwidth', fig.width = 10, fig.height = 3.9, warning = FALSE>>=

@@

exp_scatter <- 
  plot_explain_scatter(sine_poi_lime_default) +
  facet_grid(switch = "both") +
  labs(x = TeX("x_1"),
       y = TeX("x_2"),
       size = "Weight",
       shape = "",
       fill = "Random \nForest \nProbability \nfor 'blue'",
       color = "Random \nForest \nProbability \nfor 'blue'",
       linetype = "",
       title = "Explanation Scatterplot") +
  theme_bw(base_family = "Times", base_size = 10) +
  guides(shape = guide_legend(order = 1)) +
  theme(aspect.ratio = 1)

plot_grid(sine_poi_exp, exp_scatter, align = "h") 
@
\caption{(left) Visualization from the \emph{lime} R package of the LIME explanation for the \data \ prediction of interest. The bars represent the explainer model coefficients. (right) \emph{Explanation Scatterplot}. The points represent the simulated data colored by the random forest probabilities and sized by the assigned LIME weights. \hh{XXX could you make the complex prediction **much** lighter for this example? The diamond and the lines are hard to see, but should be the focus.} The prediction of interest is shown as the diamond shaped point. The explainer model indicator variables associated with $x_1$ and $x_2$ (quantile bins containing the prediction of interest) are depicted by the solid lines. The line colors represent the explainer model coefficient signs and indicate whether the feature supports a prediction of `blue' (blue lines) or supports a prediction of `red' (red lines). This figure shows that the quantile bins are not flexible enough to capture the relationship between the random forest predictions and the $x_1$ and $x_2$ values.}
\label{fig:sine-exp-scatter}
\end{figure*}

@@

\subsection{A LIME explanation}
\hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.}
To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package. The figure shows that the random forest returned a probability of \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} that the observation of interest belongs to category blue {(denoted as the ``label" in the figure).The lengths of the bars represent the explainer model coefficients associated with the indicator variables chosen by LIME as important. The color of the bar denotes the sign of the coefficient and whether the feature supports or contradicts a random forest classification of `blue'. The "explanation fit" value is the deviance ratio from the R package \emph{glmnet} \citep{simon:2011} for a ridge regression model. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014}, and it should not be used as the only metric in a model assessment.

@@

\subsection{Explanation Scatterplots} \label{exp-scatter}
\hh{XXX I think we could reduce some of the example for the specific example a bit, now that the LIME explanation is moved out, there is not that much need to repeat the general ideas.}

@@

\hh{An explanation scatterplot for the example of the previous section is shown on the right hand side of \autoref{fig:sine-exp-scatter}.}
 The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.

By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model. First consider the claim of localness. The weights remain relatively high  outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bin containing the prediction of interest. While it is still difficult to say whether or not the claim of localness has been violated, it is possible to say that the weights assigned to the simulated data do not agree with the local region assigned by the intersection of the quantile bins.

@@

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} In that respect, the diagnostics shown for the \data\ and in the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{gold}.

>>>>

# Load packages
library(caret)
library(cowplot)
library(dplyr)
library(forcats)
library(ggforce)
library(ggplot2)
library(ggpcp) #devtools::install_github("yaweige/ggpcp", build_vignettes = TRUE)
library(glmnet)
library(gower)
library(graphics)
library(gretchenalbrecht) #devtools::install_github("dicook/gretchenalbrecht")
library(latex2exp)
library(lime) #devtools::install_github("goodekat/lime")
library(limeaid) #devtools::install_github("goodekat/limeaid")
library(purrr)
library(randomForest)
library(stringr)
library(tidyr)

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(data = lime_data_gathered,
             mapping = aes(x = feature_stnd_value,
                           y = prediction,
                           size = distance_value,
                           alpha = distance_value,
                           color = distance)) +
  geom_abline(data = coefs_data %>% filter(case == "good"),
              mapping = aes(intercept = int, slope = slope),
              size = 1) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c(NA, "grey30"), guide = "none") +
  geom_point(data = prediction_of_interest_gathered,
             mapping = aes(x = feature_value,
                           y = prediction,
                           fill = color),
             color = "black",
             size = 5,
             shape = 23,
             alpha = 0.75) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(feature_stnd_value = 1.2,
                      prediction = 1.6,
                      feature = c("Feature 1", "Feature 2"),
                      slope = c(paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "good",
                                               feature == "Feature 1") %>%
                                        pull(slope) %>%
                                        round(3)),
                                paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "good",
                                               feature == "Feature 2") %>%
                                        pull(slope) %>%
                                        round(3)))),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = 4) +
  labs(x = "",
       y = "Black-Box Prediction",
       title = "Conceptual Depiction of a Faithful Local Explainer Model",
       subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
       fill = "",
       alpha = "Weight",
       size = "Weight") +
  theme_linedraw(base_family = "Times", base_size = 10) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(color = 'black', size = 10)) +
  guides(fill = guide_legend(order = 1),
         size = guide_legend(order = 2),
         alpha = guide_legend(order = 2))

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(data = lime_data_gathered,
             mapping = aes(x = feature_stnd_value,
                           y = prediction,
                           size = distance_value,
                           alpha = distance_value,
                           color = distance)) +
  geom_abline(data = coefs_data %>% filter(case == "bad"),
              mapping = aes(intercept = int, slope = slope),
              size = 1) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c("grey30", NA), guide = "none") +
  geom_point(data = prediction_of_interest_gathered,
             mapping = aes(x = feature_value,
                           y = prediction,
                           fill = color),
             color = "black",
             size = 5,
             shape = 23,
             alpha = 0.75) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(feature_stnd_value = 1.2,
                      prediction = 1.6,
                      feature = c("Feature 1", "Feature 2"),
                      slope = c(paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "bad",
                                               feature == "Feature 1") %>%
                                        pull(slope) %>%
                                        round(3)),
                                paste("Slope:",
                                      coefs_data %>%
                                        filter(case == "bad",
                                               feature == "Feature 2") %>%
                                        pull(slope) %>%
                                        round(3)))),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = 4) +
  labs(x = "",
       y = "Black-Box Prediction",
       title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
       subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
       fill = "",
       alpha = "Weight",
       size = "Weight") +
  theme_linedraw(base_family = "Times", base_size = 10) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(color = 'black', size = 10)) +
  guides(fill = guide_legend(order = 1),
         size = guide_legend(order = 2),
         alpha = guide_legend(order = 2))

@@

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two predictor variables used to fit the hypothetical model. The diamond shaped points represent a prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. \hh{In the example of \autoref{fig:concept-plot-good}}, an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of \hh{two features in the data} % features Features 1 and~2 
as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

@@

To demonstrate the visualizations, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with three features and one response variable. The features, $x_1$, $x_2$, and $x_3$, are randomly sampled from Unif(\Sexpr{min}, \Sexpr{max}), Unif(\Sexpr{min}, \Sexpr{max}), and $\mbox{N}(0,1)$ distributions, respectively. A binary response variable $y$ is  created using a rotated sine curve. In particular, let $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$ and $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$ where $\theta=-0.9$. Then $y$ is defined as
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ . %\nolabel
  \end{cases}
\end{eqnarray}
Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$.
\hh{From the plots in \autoref{sine-plot-data}}\kge{,} \hh{we see that the global relationship between response $y$ and features $x_1$ and $x_2$ is linear: the probability for label blue increases with the difference between features $x_2$ and $x_1$. Locally, the relationship between $y$ and features $x_2$ and $x_1$ varies a lot more around the line of identity. Here, the relationship is determined by the sine waves. However, the sine is a good-natured function that can be approximated well linearly} \kge{in local regions}.

@@

\hh{All runs of LIME in this paper are executed in}
a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. 
The forked version \hh{is functionally indistinguishable from Pedersen's implementation}\kge{,} \hh{but it} allows us to export internal values \hh{relevant for an assessment of the explainer}. 
The R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid} (\Sexpr{packageVersion("limeaid")}) implements diagnostic plots based on the output of the altered version of \emph{lime}.

@@

\begin{figure*}[!thp]
<<sine-exp-scatter, out.width = '\\textwidth', fig.width = 12, fig.height = 4.68, warning = FALSE>>=

@@

exp_scatter <- 
  plot_explain_scatter(sine_poi_lime_default, alpha = 0.4) +
  facet_grid(switch = "both") +
  labs(x = TeX("x_1"),
       y = TeX("x_2"),
       size = "Weight",
       shape = "",
       fill = "Random \nForest \nProbability \nfor 'blue'",
       color = "Random \nForest \nProbability \nfor 'blue'",
       linetype = "",
       title = "Explanation Scatterplot") +
  theme_bw(base_family = "Times", base_size = 10) +
  guides(shape = guide_legend(order = 1),
         size = guide_legend(nrow = 2, byrow = T)) +
  theme(aspect.ratio = 1)

plot_grid(sine_poi_exp, exp_scatter) 
@
\caption{(left) Visualization from the \emph{lime} R package of the LIME explanation for the \data \ prediction of interest. The bars represent the explainer model coefficients. (right) \emph{Explanation Scatterplot}. The points represent the simulated data colored by the random forest probabilities and sized by the assigned LIME weights. \hh{XXX could you make the complex prediction **much** lighter for this example? The diamond and the lines are hard to see, but should be the focus.} \kgc{Is that light enough? Feel free to edit the alpha value in the plot\_explain\_scatter function as you feel necessary. I also made the lines depicting the explainer model wider in limeaid, and there is an option 'line\_size' that can be adjusted.} The prediction of interest is shown as the diamond shaped point. The explainer model indicator variables associated with $x_1$ and $x_2$ (quantile bins containing the prediction of interest) are depicted by the solid lines. The line colors represent the explainer model coefficient signs and indicate whether the feature supports a prediction of `blue' (blue lines) or supports a prediction of `red' (red lines). This figure shows that the quantile bins are not flexible enough to capture the relationship between the random forest predictions and the $x_1$ and $x_2$ values.}
\label{fig:sine-exp-scatter}
\end{figure*}

@@

\subsection{A LIME explanation}
\hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.} \kgc{Okay.} To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package. The figure shows that the random forest returned a probability of \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} that the observation of interest belongs to category blue {(denoted as the ``label" in the figure).The lengths of the bars represent the explainer model coefficients associated with the indicator variables chosen by LIME as important. The color of the bar denotes the sign of the coefficient and whether the feature supports or contradicts a random forest classification of `blue'. The "explanation fit" value is the deviance ratio from the R package \emph{glmnet} \citep{simon:2011} for a ridge regression model. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014}, and it should not be used as the only metric in a model assessment.

@@

\subsection{Explanation Scatterplots} \label{exp-scatter}
\hh{XXX I think we could reduce some of the example for the specific example a bit, now that the LIME explanation is moved out, there is not that much need to repeat the general ideas.} \kgc{I commented out the text that I think you were referring to. See what you think.}

@@

\hh{An explanation scatterplot} \sout{for the} \kge{associated with the LIME explanation depicted on the left in \autoref{fig:sine-exp-scatter}} \sout{example of the previous section} \hh{is shown on the right hand side of \autoref{fig:sine-exp-scatter}.} %The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.
By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model. First consider the claim of localness. The weights remain relatively high outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bin containing the prediction of interest. While it is still difficult to say whether or not the claim of localness has been violated, it is possible to say that the weights assigned to the simulated data do not agree with the local region assigned by the intersection of the quantile bins.

@@

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} \kgc{Yes. I see what you mean now.} In that respect, the diagnostics shown for the \data\ and in the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{gold}.
